XXXXX

### 2014-12-15 paper/RDF validation requirements/RDF validation and inferencing.tex
validation and inferencing

\newenvironment{DL}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
  \begin{tabular}{l}
	\textbf{DL:} \\

@@

\section{Motivation}

\begin{itemize}
	\item mappings from RDF validation requirements (more specifically RDF constraint formulation requirements) to DL
	\item mappings from RDF validation requirements (more specifically RDF constraint formulation requirements) to generic constraints (expressed by generic constraint language)
\end{itemize}

\tb{I think we should provide mappings to DL directly in the requirements chapters and not just in 1 table at the end of the paper}

\section{Subsumption}

requirement:

\begin{itemize}
	\item R-100-SUBSUMPTION
\end{itemize}

\subsection{Description}

\begin{itemize}
	\item DL terminology: concept inclusion
\end{itemize}

\subsection{Simple Example}

\begin{DL}
Mother $\sqsubseteq$ Parent 
\end{DL}

\begin{itemize}
	\item all mothers are parents
  \item the concept Mother is subsumed by the concept Parent
\end{itemize}

\begin{gcotable}
class & Mother & - & - & Parent & $\sqsubseteq$ \\
\end{gcotable}

\subsection{Simple Example}

\begin{itemize}
	\item Jedis feel the force 
\end{itemize}

\begin{ex}
# DSP
# ---
-
\end{ex}

\begin{ex}
# OWL2
# ----
:Jedi rdfs:subClassOf :FeelingForce .
\end{ex}

\begin{ex}
# ReSh
# ----
the extension ext:extendsShape may be used
\end{ex}

\begin{ex}
# ShEx
# ----
:FeelingForce {
    :feelingForce (true) }
:Jedi {
    & :FeelingForce ,
    :attitute ('good') }
\end{ex}

\begin{ex}
# data
# ----
:Yoda 
    :feelingForce true ;
    :attitute 'good' .
\end{ex}

\begin{DL}
:Jedi $\sqsubseteq$ :FeelingForce
\end{DL}

\subsection{Complex Example}

\begin{DL}
Rich $\sqsubseteq$ $\neg$ Poor 
\end{DL}

\begin{gcotable}
class & $\neg$ Poor & - & - & Poor & $\neg$ \\
class & Rich & - & - & $\neg$ Poor & $\sqsubseteq$ \\
\end{gcotable}

\section{Class Equivalence}

requirement:

\begin{itemize}
	\item R-3-EQUIVALENT-CLASSES
\end{itemize}

\subsection{Description}

\begin{itemize}
	\item Concept equivalence asserts that two concepts have the same instances \cite{Kroetzsch2012}
  \item While synonyms are an obvious example of equivalent concepts, in practice one more
often uses concept equivalence to give a name to complex expression  \cite{Kroetzsch2012}
\item Concept equivalence is indeed subsumption  from left and right. ($A \sqsubseteq B$ and $B \sqsubseteq A$ implies $A \equiv B$ )

\end{itemize}

\subsection{Simple Example}

\begin{DL}
Person $\equiv$ Human
\end{DL}

\begin{gcotable}
class & Person & - & - & Human & $\equiv$ \\
class & Human & - & - & Person & $\equiv$ \\
\end{gcotable}

maybe better:
\er{It is not better since we should stick to what we use as syntax.}

\begin{gcotable}
class & Person & - & - & Human & $\sqsubseteq$ \\
class & Human & - & - & Person & $\sqsubseteq$ \\
\end{gcotable}
\section{Sub Object Properties}

requirements:

\begin{itemize}

	\item R-54-SUB-OBJECT-PROPERTIES
\end{itemize}

\subsection{Description}

\begin{itemize}
	\item DL terminology: role inclusion 
\end{itemize}

\subsection{Simple Example}

\begin{DL}
parentOf $\sqsubseteq$ ancestorOf 
\end{DL}

\begin{itemize}
	\item states that parentOf is a subrole of ancestorOf , i.e., every pair of individuals related by parentOf is also related by ancestorOf
\end{itemize}

\begin{gcotable}
property & parentOf & ancestorOf & - & - & $\sqsubseteq$ \\
\end{gcotable}

better:

\begin{gcotable}
property & $\top$ & parentOf & ancestorOf & $\top$ & $\sqsubseteq$ \\
\end{gcotable}

%\subsection{Simple Example}

%\begin{DL}
%sitsOn $\sqsubseteq$ touches 
%\end{DL}

\subsection{Simple Example}

\begin{DL}
hasDog $\sqsubseteq$ hasPet 
\end{DL}

\section{Sub Data Properties}

requirements:

\begin{itemize}

	\item R-54-SUB-DATA-PROPERTIES
\end{itemize}

\subsection{Description}

\begin{itemize}
	\item DL terminology: concrete domains functional roles  
\end{itemize}

\subsection{Simple Example}
\begin{DL}
 $Person \sqcap  \forall hasAge. \leq_{10} \sqsubseteq Person \sqcap \forall hasEvenAge. \leq_{10}$ 
\end{DL}

\section{Object Property Paths / Object Property Chains}

requirement:

\begin{itemize}
	\item R-55-OBJECT-PROPERTY-PATHS
\end{itemize}

\subsection{Description}

\begin{itemize}
  \item DL terminology: complex role inclusion axiom / role composition
	\item role composition can only appear on the left-hand side of complex role inclusions \cite{Kroetzsch2012}
\end{itemize}

\subsection{Simple Example}

\begin{DL}
brotherOf $\circ$ parentOf $\sqsubseteq$ uncleOf 
\end{DL}

\begin{ex}
# OWL2
# ----
:uncleOf owl:propertyChainAxiom ( :brotherOf :parentOf ) . 
\end{ex}

\begin{gcotable}
property & $\top$ & brotherOf, parentOf & uncleOf & $\top$ & $\sqsubseteq$ & -  \\
\end{gcotable}

old mapping:

\begin{gcotable}
property & uncleOf & brotherOf, parentOf & - & $\top$ & $\sqsubseteq$ \\
\end{gcotable}

\section{Disjoint Properties}

requirements:

\begin{itemize}
	\item R-9-DISJOINT-PROPERTIES
\end{itemize}

\subsection{Description}

\begin{itemize}
	\item 
\end{itemize}

\subsection{Simple Example}

\begin{DL}
Disjoint(parentOf, childOf)
\end{DL}

\begin{gcotable}
property & $\top$ & parentOf & childOf & $\top$ & $\ne$ \\
\end{gcotable}

\section{Intersection}

requirements:

\begin{itemize}
  \item R-15-CONJUNCTION-OF-CLASS-EXPRESSIONS
	\item R-16-CONJUNCTION-OF-DATA-RANGES
\end{itemize}

\subsection{Description}

\begin{itemize}
	\item composition / conjunction
\end{itemize}

DLs allow new concepts and roles to be
built using a variety of different constructors. We distinguish concept and role 
constructors depending on whether concept or role expressions are constructed. In the case of
concepts, one can further separate basic Boolean constructors, role restrictions and nominals/enumerations \cite{Kroetzsch2012}.

Boolean concept constructors provide basic boolean operations that are closely related to
the familiar operations of intersection, union and complement of sets, or to conjunction,
disjunction and negation of logical expressions \cite{Kroetzsch2012}.

\begin{DL}
Mother $\equiv$ Female $\sqcap$ Parent
\end{DL}

concept inclusions allow us to state that all mothers are female and that
all mothers are parents, but what we really mean is that mothers are exactly the female
parents. DLs support such statements by allowing us to form complex concepts such as
the intersection (also called conjunction)
which denotes the set of individuals that are both female and parents. A complex concept
can be used in axioms in exactly the same way as an atomic concept, e.g., in the
equivalence Mother $\equiv$ Female $\sqcap$ Parent .

\subsection{Simple Example}

\begin{DL}
Female $\sqcap$ Parent
\end{DL}

\begin{itemize}
	\item complex concept of all individuals which are of the concept Female and of the concept Parent
\end{itemize}

\begin{gcotable}
class & Female $\sqcap$ Parent & - & - & Female, Parent & $\sqcap$ \\
\end{gcotable}

\begin{DL}
Mother $\equiv$ Female $\sqcap$ Parent
\end{DL}

\begin{gcotable}
class & Mother & - & - & Female, Parent & $\sqcap$ \\
\end{gcotable}

\section{Disjunction}

requirements:

\begin{itemize}
	\item R-17-DISJUNCTION-OF-CLASS-EXPRESSIONS
	\item R-18-DISJUNCTION-OF-DATA-RANGES
\end{itemize}

\subsection{Description}

\begin{itemize}
	\item union / inclusive OR
\end{itemize}

\subsection{Simple Example}

\begin{DL}
Father $\sqcup$ Mother $\sqcup$ Child
\end{DL}

\begin{gcotable}
class & complex concept & - & - & Father, Mother, Child & $\sqcup$ \\
\end{gcotable}

\subsection{Simple Example}

\begin{DL}
Parent $\equiv$ Father $\sqcup$ Mother
\end{DL}

\begin{gcotable}
class & Parent & - & - & Father, Mother & $\sqcup$ \\
\end{gcotable}

\section{Negation}

requirements:

\begin{itemize}
  \item R-19-NEGATION-OF-CLASS-EXPRESSIONS
	\item R-20-NEGATION-OF-DATA-RANGES
\end{itemize}

\subsection{Description}

\begin{itemize}
	\item complement
\end{itemize}

\subsection{Simple Example}

\begin{DL}
$ \neg Married $ \\ 
\end{DL}

\begin{itemize}
	\item set of all individuals that are not married
\end{itemize}

\begin{gcotable}
class & $\neg$ Married & - & - & Married & $\neg$ \\
\end{gcotable}

\subsection{Complex Example}

\begin{DL}
Female $\sqcap$ $\neg$ Married \\
\end{DL}

\begin{gcotable}
class & $\neg$ Married & - & - & Married & $\neg$ \\
class & Female $\sqcap$ $\neg$ Married & - & - & Female, $\neg$ Married & $\sqcap$ \\
\end{gcotable}

\section{Disjoint Classes}

requirements:

\begin{itemize}
	\item R-7-DISJOINT-CLASSES
\end{itemize}

\subsection{Description}

\begin{itemize}
	\item 
\end{itemize}

\subsection{Simple Example}

\begin{DL}
Male $\sqcap$ Female $\sqsubseteq$ $\perp$
\end{DL}

\begin{gcotable}
class & Male $\sqcap$ Female & - & - & Male, Female & $\sqcap$  \\
class & $\top$ & - & - & Male $\sqcap$ Female, $\perp$ & $\sqsubseteq$ \\
\end{gcotable}

syntactic sugar:

\begin{gcotable}
class & Male & - & - & Female & $\ne$ \\
class & Female & - & - & Male & $\ne$ \\
\end{gcotable}

\section{Existential Quantification on Properties}

requirements:

\begin{itemize}
	\item R-86-EXISTENTIAL-QUANTIFICATION-ON-PROPERTIES
\end{itemize}

\subsection{Description}

\begin{itemize}
	\item DL termonology: existential restriction
	\item An existential class expression ObjectSomeValuesFrom( OPE CE ) consists of an object property expression OPE and a class expression CE, and it contains all those individuals that are connected by OPE to an individual that is an instance of CE. Such a class expression can be seen as a syntactic shortcut for the class expression ObjectMinCardinality( 1 OPE CE ).
	\item An existential class expression DataSomeValuesFrom( DPE1 ... DPEn DR ) consists of n data property expressions DPEi, 1 $\leq$ i $\leq$ n, and a data range DR whose arity MUST be n. Such a class expression contains all those individuals that are connected by DPEi to literals lti, 1 $\leq$ i $\leq$ n, such that the tuple ( lt1 , ..., ltn ) is in DR. A class expression of the form DataSomeValuesFrom( DPE DR ) can be seen as a syntactic shortcut for the class expression DataMinCardinality( 1 DPE DR ). 
\end{itemize}

\subsection{Simple Example}

\begin{DL}
$\exists$ parentOf . $\top$
\end{DL}

complex concept that describes the set
of individuals that are parents of at least one individual (instance of $\top$).

\begin{gcotable}
property & $\exists$ parentOf . $\top$ & parentOf & - & $\top$ & $\exists$ \\
\end{gcotable}

\subsection{Simple Example}

\begin{DL}
$\exists$ parentOf . Female
\end{DL}

the complex concept describes those individuals that are parents of at least one
female individual, i.e., those that have a daughter.

\begin{gcotable}
property & $\exists$ parentOf . Female & parentOf & - & Female & $\exists$ \\
\end{gcotable}

\subsection{Simple Example}

\begin{DL}
Parent $\equiv$ $\exists$ parentOf . $\top$
\end{DL}

\begin{gcotable}
property & Parent & parentOf & - & $\top$ & $\exists$ \\
\end{gcotable}

\section{Universal Quantification on Properties}

requirements:

\begin{itemize}
	\item R-91-UNIVERSAL-QUANTIFICATION-ON-PROPERTIES
\end{itemize}

\subsection{Description}

\begin{itemize}
	\item DL termonology: universal restriction
\end{itemize}

\tb{definition}

\subsection{Simple Example}

\begin{DL}
$\forall$ parentOf . Female
\end{DL}

\begin{itemize}
	\item set of individuals all of whose children are female
  \item also includes those that have no children at all
\end{itemize}

\begin{gcotable}
property & $\forall$ parentOf . Female & parentOf & - & Female & $\forall$ \\
\end{gcotable}

@@

requirements:

\begin{itemize}
	\item R-25-OBJECT-PROPERTY-DOMAIN
	\item R-26-DATA-PROPERTY-DOMAIN
\end{itemize}

\subsection{Description}

\begin{itemize}
	\item restrict domain of object and data properties
	\item DL terminology: domain restrictions on roles
	\item The language needs to include an easy-to-use vocabulary to declare that a given property is associated with a class, e.g. to populate input forms with appropriate widgets but also constraint checking. In OO terms this is the declaration of a member, field, attribute or association. 
\end{itemize}

\tb{definition}

\subsection{Simple Example}

\begin{DL}
$\exists$ sonOf . $\top$ $\sqsubseteq$ Male 
\end{DL}

\begin{itemize}
	\item restrict the domain of sonOf to male individual
\end{itemize}

\begin{gcotable}
property & $\top$ & sonOf & - & Male & domain \\
\end{gcotable}

\section{Property Range}

requirements:

\begin{itemize}
	\item R-28-OBJECT-PROPERTY-RANGE
	\item R-35-DATA-PROPERTY-RANGE
\end{itemize}

\subsection{Description}

\begin{itemize}
	\item restrict range of object and data properties
	\item DL terminology: range restrictions on roles
\end{itemize}

\tb{definition}

\subsection{Simple Example}

\begin{DL}
$\top$ $\sqsubseteq$ $\forall$ sonOf . Parent 
\end{DL}

\begin{itemize}
	\item restrict the range of sonOf to parents
\end{itemize}

\begin{ex}
# OWL 2
# -----
sonOf rdfs:range Parent . 
\end{ex}

\begin{ex}
# DSP
# ---
:hasDogRange
        a dsp:DescriptionTemplate ; 
        dsp:resourceClass owl:Thing ; 
        dsp:statementTemplate [
            a dsp:NonLiteralStatementTemplate ;
            dsp:property sonOf ; 
            dsp:nonLiteralConstraint [ 
                a dsp:NonLiteralConstraint ;
                dsp:valueClass Parent ] ] .
\end{ex}

\begin{gcotable}
property & $\top$ & sonOf & - & Parent & range \\
\end{gcotable}

\section{Class-Specific Property Range}

requirements:

\begin{itemize}
	\item R-29-CLASS-SPECIFIC-RANGE-OF-RDF-OBJECTS
	\item R-36-CLASS-SPECIFIC-RANGE-OF-RDF-LITERALS
\end{itemize}

\subsection{Description}

\begin{itemize}
  \item restrict range of object and data properties for individuals within specific context
	\item context: e.g. class
	\item The values of each member property of a class may be limited by their value type, such as xsd:string or ex:Person. 
\end{itemize}

\subsection{Simple Example}

\begin{itemize}
	\item property 'sonOf'
	\item context: class 'Man'
	\item range: class 'Parent'
	\item $\rightarrow$ individuals of the class 'Man' can only have 'sonOf' relationships to indivuduals of the class 'Parent'
\end{itemize}

\section{Minimum Unqualified Cardinality Restrictions on Properties}

requirements:

\begin{itemize}
	\item R-81-MINIMUM-UNQUALIFIED-CARDINALITY-ON-PROPERTIES
	\item R-211-CARDINALITY-CONSTRAINTS
\end{itemize}

description:

\begin{itemize}
	\item A minimum cardinality expression ObjectMinCardinality( n OPE CE ) consists of a nonnegative integer n, an object property expression OPE, and a class expression CE, and it contains all those individuals that are connected by OPE to at least n different individuals that are instances of CE. If CE is missing, it is taken to be owl:Thing.
	\item A minimum cardinality expression DataMinCardinality( n DPE DR ) consists of a nonnegative integer n, a data property expression DPE, and a unary data range DR, and it contains all those individuals that are connected by DPE to at least n different literals in DR. If DR is not present, it is taken to be rdfs:Literal. 
	\item unqualified: CE respective DR is NOT stated. 
\end{itemize}

\subsection{Simple Example}

\begin{ex}
# ShEx
# ----
:Jedi {
    :attitude ('good') }
:JediStudent {
    & :Jedi ,
    :studentOf {}{1} }
:JediMaster {
    & :Jedi ,
    :mentorOf {}{1,2} }
:SuperJediMaster {
    & :Jedi ,
    :mentorOf {}{3,} }
\end{ex}

\begin{itemize}
	\item Jedis have the attitude 'good'
	\item Jedi students are students of exactly 1 resource
	\item Jedi masters are mentoring at least 1 and at most 2 resources
	\item Super Jedi masters are mentoring at least 3 resources  
\end{itemize}

\begin{ex}
# data
# ----
:Yoda 
    :attitude 'good' ;
    :mentorOf :MaceWindu , :Obi-Wan , :Luke .
:MaceWindu
    :attitude 'good' ;
    :studentOf :Yoda .
:Obi-Wan 
    :attitude 'good' ;
    :studentOf :Yoda ;
    :mentorOf :Anakin .
:Anakin
    :attitude 'good' ; 
    :studentOf :Obi-Wan .
:Luke
    :attitude 'good' ;
    :studentOf :Yoda .
\end{ex}

\begin{ex}
# Individuals matching the ’:SuperJediMaster’ data shape
# ------------------------------------------------------
:Yoda 
\end{ex}

\tb{ToDO: add DL}

\section{Minimum Qualified Cardinality Restrictions on Properties}

requirements:

\begin{itemize}
	\item R-75-MINIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES
  \item R-211-CARDINALITY-CONSTRAINTS
\end{itemize}

description:

\begin{itemize}
	\item A minimum cardinality expression ObjectMinCardinality( n OPE CE ) consists of a nonnegative integer n, an object property expression OPE, and a class expression CE, and it contains all those individuals that are connected by OPE to at least n different individuals that are instances of CE. If CE is missing, it is taken to be owl:Thing.
	\item A minimum cardinality expression DataMinCardinality( n DPE DR ) consists of a nonnegative integer n, a data property expression DPE, and a unary data range DR, and it contains all those individuals that are connected by DPE to at least n different literals in DR. If DR is not present, it is taken to be rdfs:Literal. 
	\item qualified: CE respective DR is stated.
\end{itemize}

\subsection{Simple Example}

\begin{DL}
$\geq$ 2 childOf . Parent
\end{DL}

\tb{ToDO: check DL}

\begin{itemize}
	\item set of individuals that are children of at least two parents
\end{itemize}

\begin{ex}
# OWL 2
# -----
owl:Thing
    a owl:Restriction ;
    owl:minQualifiedCardinality "2"^^xsd:nonNegativeInteger ;
    owl:onProperty childOf ;
    owl:onClass Parent .
\end{ex}

\begin{gcotable}
property & $\geq$ 2 childOf . Parent & childOf & -  & Parent & $\geq$ & 2 \\
\end{gcotable}

\subsection{Simple Example}

\begin{DL}
foaf:Person $\equiv$ $\geq$ 2 hasName.xsd:string
\end{DL}

\tb{ToDO: check DL}

\begin{gcotable}
property & foaf:Person & hasName & - & xsd:string & $\geq$ & 2 \\
\end{gcotable}

\subsection{Simple Example}

\tb{ToDO: OWL 2 DL}

\begin{ex}
# ShEx
# ----
:Jedi {
    :attitude ('good') }
:JediStudent {
    & :Jedi ,
    :studentOf @:Jedi{1} }
:JediMaster {
    & :Jedi ,
    :mentorOf @:Jedi{1,2} }
:SuperJediMaster {
    & :Jedi ,
    :mentorOf @:Jedi{3,} }
\end{ex}

\begin{itemize}
	\item Jedis have the attitude 'good'
	\item Jedi students are students of exactly 1 Jedi
	\item Jedi masters are mentoring at least 1 and at most 2 Jedis
	\item Super Jedi masters are mentoring at least 3 Jedis
\end{itemize}

\begin{ex}
# data
# ----
:Yoda 
    :attitude 'good' ;
    :mentorOf :MaceWindu , :Obi-Wan , :Luke .
:MaceWindu
    :attitude 'good' ;
    :studentOf :Yoda .
:Obi-Wan 
    :attitude 'good' ;
    :studentOf :Yoda ;
    :mentorOf :Anakin .
:Anakin
    :attitude 'good' ; 
    :studentOf :Obi-Wan .
:Luke
    :attitude 'good' ;
    :studentOf :Yoda .
\end{ex}

\begin{ex}
# Individuals matching the ’:SuperJediMaster’ data shape
# ------------------------------------------------------
:Yoda

# Individuals matching the ’:JediMaster’ data shape
# -------------------------------------------------
:Obi-Wan
\end{ex}

\tb{ToDO: add DL}

\subsection{Complex Example}

\begin{DL}
Father2Daughters $\equiv$ Man $\sqcap$ ($\geq$ 2 hasChild.Woman)
\end{DL}

\tb{ToDO: check DL}

\begin{gcotable}
property & $\geq$ 2 hasChild.Woman & hasChild & - & Woman & $\geq$ & 2 \\
class & Father2Daughters & - & - & Man, $\geq$ 2 hasChild.Woman & $\sqcap$ & - \\
\end{gcotable}

\section{Maximum Unqualified Cardinality Restrictions on Properties}

requirements:

\begin{itemize}
	\item R-82-MAXIMUM-UNQUALIFIED-CARDINALITY-ON-PROPERTIES
	\item R-211-CARDINALITY-CONSTRAINTS
\end{itemize}

description:

\begin{itemize}
	\item A maximum cardinality expression ObjectMaxCardinality( n OPE CE ) consists of a nonnegative integer n, an object property expression OPE, and a class expression CE, and it contains all those individuals that are connected by OPE to at most n different individuals that are instances of CE. If CE is missing, it is taken to be owl:Thing.
	\item A maximum cardinality expression DataMaxCardinality( n DPE DR ) consists of a nonnegative integer n, a data property expression DPE, and a unary data range DR, and it contains all those individuals that are connected by DPE to at most n different literals in DR. If DR is not present, it is taken to be rdfs:Literal. 
	\item unqualified: CE respective DR is NOT stated. 
\end{itemize}

\subsection{Simple Example}

\begin{ex}
# ShEx
# ----
:Jedi {
    :attitude ('good') }
:JediStudent {
    & :Jedi ,
    :studentOf {}{1} }
:JediMaster {
    & :Jedi ,
    :mentorOf {}{1,2} }
:SuperJediMaster {
    & :Jedi ,
    :mentorOf {}{3,} }
\end{ex}

\begin{itemize}
	\item Jedis have the attitude 'good'
	\item Jedi students are students of exactly 1 resource
	\item Jedi masters are mentoring at least 1 and at most 2 resources
	\item Super Jedi masters are mentoring at least 3 resources  
\end{itemize}

\begin{ex}
# data
# ----
:Yoda 
    :attitude 'good' ;
    :mentorOf :MaceWindu , :Obi-Wan , :Luke .
:MaceWindu
    :attitude 'good' ;
    :studentOf :Yoda .
:Obi-Wan 
    :attitude 'good' ;
    :studentOf :Yoda ;
    :mentorOf :Anakin .
:Anakin
    :attitude 'good' ; 
    :studentOf :Obi-Wan .
:Luke
    :attitude 'good' ;
    :studentOf :Yoda .
\end{ex}

\begin{ex}
# Individuals matching the ’:JediMaster’ data shape
# -------------------------------------------------
:Obi-Wan
\end{ex}

\tb{ToDO: add DL}

\section{Maximum Qualified Cardinality Restrictions on Properties}

requirements:

\begin{itemize}
	\item R-76-MAXIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES
	\item R-211-CARDINALITY-CONSTRAINTS
\end{itemize}

description:

\begin{itemize}
	\item A maximum cardinality expression ObjectMaxCardinality( n OPE CE ) consists of a nonnegative integer n, an object property expression OPE, and a class expression CE, and it contains all those individuals that are connected by OPE to at most n different individuals that are instances of CE. If CE is missing, it is taken to be owl:Thing.
	\item A maximum cardinality expression DataMaxCardinality( n DPE DR ) consists of a nonnegative integer n, a data property expression DPE, and a unary data range DR, and it contains all those individuals that are connected by DPE to at most n different literals in DR. If DR is not present, it is taken to be rdfs:Literal. 
	\item qualified: CE respective DR is stated. 
\end{itemize}

\subsection{Simple Example}

\begin{DL}
$\leq$  2 childOf . Parent
\end{DL}

\tb{ToDO: check DL}

\begin{itemize}
	\item set of individuals that are children of at most two parents
\end{itemize}

\begin{gcotable}
property & $\leq$  2 childOf . Parent & childOf & - & Parent & $\leq$ & 2 \\
\end{gcotable}

\subsection{Simple Example}

\tb{ToDO: OWL 2 DL}

\begin{ex}
# ShEx
# ----
:Jedi {
    :attitude ('good') }
:JediStudent {
    & :Jedi ,
    :studentOf @:Jedi{1} }
:JediMaster {
    & :Jedi ,
    :mentorOf @:Jedi{1,2} }
:SuperJediMaster {
    & :Jedi ,
    :mentorOf @:Jedi{3,} }
\end{ex}

\begin{itemize}
	\item Jedis have the attitude 'good'
	\item Jedi students are students of exactly 1 Jedi
	\item Jedi masters are mentoring at least 1 and at most 2 Jedis
	\item Super Jedi masters are mentoring at least 3 Jedis
\end{itemize}

\begin{ex}
# data
# ----
:Yoda 
    :attitude 'good' ;
    :mentorOf :MaceWindu , :Obi-Wan , :Luke .
:MaceWindu
    :attitude 'good' ;
    :studentOf :Yoda .
:Obi-Wan 
    :attitude 'good' ;
    :studentOf :Yoda ;
    :mentorOf :Anakin .
:Anakin
    :attitude 'good' ; 
    :studentOf :Obi-Wan .
:Luke
    :attitude 'good' ;
    :studentOf :Yoda .
\end{ex}

\begin{ex}
# Individuals matching the ’:JediMaster’ data shape
# -------------------------------------------------
:Obi-Wan
\end{ex}

\tb{ToDO: add DL}

\section{Exact Unqualified Cardinality Restrictions on Properties}

requirements:

\begin{itemize}
	\item R-80-EXACT-UNQUALIFIED-CARDINALITY-ON-PROPERTIES
	\item R-211-CARDINALITY-CONSTRAINTS
\end{itemize}

description:

\begin{itemize}
	\item An exact cardinality expression ObjectExactCardinality( n OPE CE ) consists of a nonnegative integer n, an object property expression OPE, and a class expression CE, and it contains all those individuals that are connected by OPE to exactly n different individuals that are instances of CE. If CE is missing, it is taken to be owl:Thing. Such an expression is actually equivalent to the expression ObjectIntersectionOf( ObjectMinCardinality( n OPE CE ) ObjectMaxCardinality( n OPE CE ) ).
	\item An exact cardinality expression DataExactCardinality( n DPE DR ) consists of a nonnegative integer n, a data property expression DPE, and a unary data range DR, and it contains all those individuals that are connected by DPE to exactly n different literals in DR. If DR is not present, it is taken to be rdfs:Literal. 
	\item unqualified: CE respective DR is NOT stated. 
\end{itemize}

\subsection{Simple Example}

\tb{ToDo. OWL 2 DL}

\begin{ex}
# ShEx
# ----
:JediStudent {
    :studentOf {}{1} }
\end{ex}

\begin{ex}
# ReSh
# ----
:JediStudent a rs:ResourceShape ;
    rs:property [
        rs:name "studentOf" ;
        rs:propertyDefinition :studentOf ;
        rs:valueShape [ a rs:ResourceShape] ;
        rs:occurs rs:Exactly-one ; ] .
\end{ex}

\begin{itemize}
	\item Jedis have the attitude 'good'
	\item Jedi students are students of exactly 1 resource
\end{itemize}

\begin{ex}
# data
# ----
:Yoda 
    :attitude 'good' ;
    :mentorOf :MaceWindu , :Obi-Wan , :Luke .
:MaceWindu
    :attitude 'good' ;
    :studentOf :Yoda .
:Obi-Wan 
    :attitude 'good' ;
    :studentOf :Yoda ;
    :mentorOf :Anakin .
:Anakin
    :attitude 'good' ; 
    :studentOf :Obi-Wan .
:Luke
    :attitude 'good' ;
    :studentOf :Yoda .
\end{ex}

\begin{ex}
# Individuals matching the ’:JediStudent’ data shape
# --------------------------------------------------
:MaceWindu :Obi-Wan :Anakin :Luke
\end{ex}

\tb{ToDO: add DL}

\section{Exact Qualified Cardinality Restrictions on Properties}

requirements:

\begin{itemize}
	\item R-74-EXACT-QUALIFIED-CARDINALITY-ON-PROPERTIES
	\item R-211-CARDINALITY-CONSTRAINTS
\end{itemize}

description:

\begin{itemize}
	\item An exact cardinality expression ObjectExactCardinality( n OPE CE ) consists of a nonnegative integer n, an object property expression OPE, and a class expression CE, and it contains all those individuals that are connected by OPE to exactly n different individuals that are instances of CE. If CE is missing, it is taken to be owl:Thing. Such an expression is actually equivalent to the expression ObjectIntersectionOf( ObjectMinCardinality( n OPE CE ) ObjectMaxCardinality( n OPE CE ) ).
	\item An exact cardinality expression DataExactCardinality( n DPE DR ) consists of a nonnegative integer n, a data property expression DPE, and a unary data range DR, and it contains all those individuals that are connected by DPE to exactly n different literals in DR. If DR is not present, it is taken to be rdfs:Literal. 
	\item qualified: CE respective DR is stated. 
\end{itemize}

\subsection{Simple Example}

\tb{ToDO: OWL 2 DL}

\begin{ex}
# ShEx
# ----
:Jedi {
    :attitude ('good') }
:JediStudent {
    :studentOf @:Jedi{1} }
\end{ex}

\begin{ex}
# ReSh
# ----
:Jedi a rs:ResourceShape ;
    rs:property [
        rs:name "attitude" ;
        rs:propertyDefinition :attitude ;
        rs:allowedValue "good" ;
        rs:occurs rs:Exactly-one ;
    ] .
:JediStudent a rs:ResourceShape ;
    rs:property [
        rs:name "studentOf" ;
        rs:propertyDefinition :studentOf ;
        rs:valueShape :Jedi ;
        rs:occurs rs:Exactly-one ;
    ] .
\end{ex}

\begin{itemize}
	\item Jedis have the attitude 'good'
	\item Jedi students are students of exactly 1 Jedi
\end{itemize}

\begin{ex}
# data
# ----
:Yoda 
    :attitude 'good' ;
    :mentorOf :MaceWindu , :Obi-Wan , :Luke .
:MaceWindu
    :attitude 'good' ;
    :studentOf :Yoda .
:Obi-Wan 
    :attitude 'good' ;
    :studentOf :Yoda ;
    :mentorOf :Anakin .
:Anakin
    :attitude 'good' ; 
    :studentOf :Obi-Wan .
:Luke
    :attitude 'good' ;
    :studentOf :Yoda .
\end{ex}

\begin{ex}
# Individuals matching the ’:JediStudent’ data shape
# --------------------------------------------------
:MaceWindu :Obi-Wan :Anakin :Luke
\end{ex}

\tb{ToDO: add DL}

\subsection{Complex Example}

\begin{DL}
Person $\sqsubseteq$ $\geq$ 2 childOf . Parent $\sqcap$ $\leq$  2 childOf . Parent \\
\end{DL}

\tb{ToDO: check DL}

\begin{itemize}
	\item every person is a child of exactly two parents
\end{itemize}

\begin{gcotable}
property & $\geq$ 2 childOf . Parent & childOf & - & Parent & $\geq$ & 2 \\
property & $\leq$ 2 childOf . Parent & childOf & - & Parent & $\leq$ & 2 \\
class & Person & - & - & $\geq$ 2 childOf . Parent, $\leq$ 2 childOf . Parent & $\sqcap$ & - \\
\end{gcotable}

syntactic sugar:

\begin{gcotable}
property & Person & childOf & - & Parent & = & 2 \\
\end{gcotable}

\section{Inverse Object Properties}

requirement:

\begin{itemize}
	\item R-56-INVERSE-OBJECT-PROPERTIES
\end{itemize}

\subsection{Description}

\begin{itemize}
	\item In many cases properties are used bi-directionally and then accessed in the inverse direction, e.g. parent $\equiv$ child$^{-}$. There should be a way to declare value type, cardinality etc of those inverse relations without having to declare a new property URI. 
	\item The object property expression OPE1 is an inverse of the object property expression OPE2. Thus, if an individual x is connected by OPE1 to an individual y, then y is also connected by OPE2 to x, and vice versa.
\end{itemize}

\subsection{Simple Example}

\begin{DL}
parentOf $\equiv$ $childOf^{-}$ \\
\end{DL}

\begin{gcotable}
property & parentOf & childOf & - & $\top$ & $inverse role (^{-})$ \\
\end{gcotable}

consistent with ontology:

\begin{gcotable}
property & $\top$ & parentOf & childOf & $\top$ & $inverse role (^{-})$ \\
\end{gcotable}

\section{Transitive Object Properties}

requirements:

\begin{itemize}
	\item R-63-TRANSITIVE-OBJECT-PROPERTIES
\end{itemize}

\subsection{Description}

\begin{itemize}
	\item Transitivity is a special form of complex role inclusion
	\item An object property transitivity axiom TransitiveObjectProperty( OPE ) states that the object property expression OPE is transitive — that is, if an individual x is connected by OPE to an individual y that is connected by OPE to an individual z, then x is also connected by OPE to z.
\end{itemize}

\subsection{Simple Example}

\begin{DL}
ancestorOf $\circ$ ancestorOf $\sqsubseteq$ ancestorOf
\end{DL}

syntactic sugar:

\begin{gcotable}
property & ancesterOf & ancestorOf, ancestorOf & - & $\top$ & transitive role \\
\end{gcotable}

syntactic sugar (ontology conform):

\begin{gcotable}
property & $\top$ & ancestorOf, ancestorOf & ancestorOf & $\top$ & transitive role \\
\end{gcotable}

without syntactic sugar (ontology conform):

>>>>

\newenvironment{DL}{
  \sffamily
  \vspace{0.3cm}
  \begin{tabular}{l}

@@

@@

The property domain constraint

\begin{DL}
$\exists$ :studentOf . $\top$ $\sqsubseteq$ :JediStudent 
\end{DL}

restricts that individuals having :studentOf relationships must be Jedi students.
This property domain constraint can also be expressed by OWL 2 QL:

\begin{ex}
:studentOf rdfs:domain :JediStudent .
\end{ex}

Without reasoning, the data \ms{:Anakin :studentOf :Obi-Wan} is invalid and causes a constraint violation, as it is not explicitly stated that \ms{:Anakin} is assigned to the class \ms{:Jedi}. 
When inferencing is performed before validating, the class assignment \ms{:Anakin rdf:type :Jedi} is inferred which prevents the constraint violation to be raised.

\tb{complexity with OWL 2 QL inferencing?}

\tb{complexity without OWL 2 QL inferencing?}
